Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

### 2024-11-15 analysis/paper/article.qmd
confirm focus on US job ads only

# explore over time
fig_prop_by_job_title_per_year <- 
jobdata %>%
  group_by(year_ad_posted) %>%
  count(job_title_simple) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot() +
  aes(year_ad_posted,
      n,
      group = job_title_simple,
      fill = job_title_simple) +
 # geom_line(linewidth = 2) +
  geom_col() +
 theme_minimal(base_size = base_size) +
  xlab("") +
  ylab("Number of\ntenure track job ads") +
  scale_fill_brewer(palette = "Dark2") +
  guides(fill = guide_legend(nrow=2,
                               byrow=TRUE,
                               "Job title",
                                keywidth=0.1,
                 keyheight=0.1,
                 default.unit="inch")) +
    theme(legend.position = c(0.5, 0.85)) +
  guides(x = "none")

@@

We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023, and focus our analysis on the 431 ads for positions at US universities. @fig-show-basic-plots shows the count of ads for each year. Assistant Professor jobs are consistently the most common title and rank, and open rank or full professor are the least frequent. The ratio of tenure-track to non-tenure track positions is generally well above one. Only 2013-2014 had more non-tenure track positions than tenure track, which was followed by an upward trend peaking at 2018-2019 and the declining again into the present.

@@

state_to_st <- function(x){
       c(state.abb, 'DC')[match(x, c(state.name, 'District of Columbia'))]
}

@@

library(ggplot2)
library(fiftystater)
library(tidyverse)
library(ggrepel)

>>>>

# explore over time
fig_prop_by_job_title_per_year <- 
jobdata %>%
  group_by(year_ad_posted) %>%
  count(job_title_simple) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot() +
  aes(year_ad_posted,
      n,
      group = job_title_simple,
      fill = job_title_simple) +
 # geom_line(linewidth = 2) +
  geom_col() +
 theme_minimal(base_size = base_size) +
  xlab("") +
  ylab("Number of\ntenure track job ads") +
  scale_fill_brewer(palette = "Dark2") +
  guides(fill = guide_legend(nrow=2,
                               byrow=TRUE,
                               "Job title",
                                keywidth=0.1,
                 keyheight=0.1,
                 default.unit="inch")) +
    theme(legend.position = c(0.5, 0.85)) +
  guides(x = "none") +
  ylim(0,100)

```{r}
# get US job ads only 
state_to_st <- function(x){
       c(state.abb, 'DC')[match(x, c(state.name, 'District of Columbia'))]
}

jobdata_us_only <- 
jobdata %>% 
  mutate(state_country = regmatches(name_of_hiring_university,
                                    gregexpr( "(?<=\\().+?(?=\\))",
                                              name_of_hiring_university,
                                              perl = T))) %>%
  unnest(state_country) %>% 
  filter(state_country %in% state.abb)

fig_prop_by_job_title_per_year_us <- 
jobdata_us_only %>%
  group_by(year_ad_posted) %>%
  count(job_title_simple) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot() +
  aes(year_ad_posted,
      n,
      group = job_title_simple,
      fill = job_title_simple) +
 # geom_line(linewidth = 2) +
  geom_col() +
 theme_minimal(base_size = base_size) +
  xlab("") +
  ylab("Number of\ntenure track job ads") +
  scale_fill_brewer(palette = "Dark2") +
  guides(fill = guide_legend(nrow=2,
                               byrow=TRUE,
                               "Job title",
                                keywidth=0.1,
                 keyheight=0.1,
                 default.unit="inch")) +
    theme(legend.position = c(0.5, 0.85)) +
  guides(x = "none") +
  ylim(0,100)

total_number_of_us_ads_in_our_sample <- nrow(jobdata_us_only)

# all of our analysis is on the US jobs ads only
jobdata <- jobdata_us_only

@@

We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023, and focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities. @fig-show-basic-plots shows the count of ads for each year. Assistant Professor jobs are consistently the most common title and rank, and open rank or full professor are the least frequent. The ratio of tenure-track to non-tenure track positions is generally well above one. Only 2013-2014 had more non-tenure track positions than tenure track, which was followed by an upward trend peaking at 2018-2019 and the declining again into the present.

@@

@@

library(ggplot2)
library(fiftystater)
# install.packages("devtools")
# devtools::install_github("wmurphyrd/fiftystater")
library(tidyverse)
library(ggrepel)
